---
title: "Direction Dependence Analysis in R: A Tutorial on the dda Package"
shorttitle: "DDA in R"
author:
  - name: "Megan J. Hirni"
    affiliation: "1"
  - name: "Wolfgang Wiedermann"
    affiliation: "1"
affiliation:
  - id: "1"
    institution: "Department of Educational, School, and Counseling Psychology, University of Missouri"
abstract: |
  [~200 words. Problem: y ~ x and x ~ y are indistinguishable under the normal
  linear model. Resolution: higher-order moment asymmetries. Contribution: the
  dda package. Payoff: worked example + decision guidance + reproducible code.]
keywords: ["direction dependence", "causal discovery", "non-normality",
           "higher-order moments", "R package", "bootstrap aggregation"]
bibliography: dda-tutorial.bib
csl: apa.csl
output:
  bookdown::pdf_document2:
    keep_tex: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  cache   = TRUE,      # DDA bootstraps are slow; cache aggressively
  comment = "#>",
  fig.align = "center",
  fig.width = 6.5, fig.height = 4
)
set.seed(20260830)     # report this seed in the Open Practices statement
library(dda)
```

<!--
HOW TO USE THIS SKELETON
- Mirrors MBR_dda_tutorial_outline.docx section for section.
- Every chunk below is a placeholder; replace `dat`, `x`, `y`, `z` with the real
  variables. Nothing here should ship with simulated data unless Section 4.1
  says so explicitly.
- Set B >= 500 for anything reported in the manuscript. Keep B small only while
  drafting, and bump it before the final knit.
- Confirm MBR's current article type, length ceiling, and abstract limit before
  writing to a budget.
-->

# Introduction

## The directionality problem in behavioral research

<!-- Substantive example carried through the whole paper. Formal statement of
     the indistinguishability of the two models under joint normality. What the
     field does instead, and where each alternative is unavailable. -->

## What non-normality buys you

<!-- Independence of predictor and residual under the correct model only.
     Relation to LiNGAM/ICA; what DDA adds. Boundary conditions stated up front:
     linearity, no confounding of the focal pair, sufficient non-normality. -->

## Aims and scope

<!-- Three aims; what this paper is not; roadmap. -->

# Statistical Background

## Notation and the competing models

## Variable-distribution tests

## Residual-distribution tests

## Independence tests

## Conditional DDA

## Robust estimation and bagging

# The dda Package

## Installation and design

```{r install, eval = FALSE}
install.packages("dda")                      # CRAN
remotes::install_github("wwiedermann/dda")   # development version
library(dda)
```

## Function overview

```{r tab-overview, echo = FALSE}
# TABLE 1: Research question | Function | Central arguments | Output
# Build with knitr::kable() or kableExtra so it renders in the PDF.
```

## Computational cost and choosing B

# Worked Example

## Data and substantive question

```{r data}
# dat <- read.csv("data/<real-open-dataset>.csv")
# Report n, descriptives, skewness and kurtosis; check the non-normality
# precondition explicitly before proceeding.
```

## Fitting the target model

```{r target-model}
# m <- lm(y ~ x + covariate1 + covariate2, data = dat)
# summary(m)
```

## Variable distributions

```{r vardist}
# vd <- dda.vardist(m, pred = "x", data = dat, B = 500)
# print(vd)
```

## Residual distributions

```{r resdist}
# rd <- dda.resdist(m, pred = "x", data = dat, B = 500, prob.trans = TRUE)
# print(rd)
```

## Independence properties

```{r indep}
# ind <- dda.indep(m, pred = "x", data = dat, nlfun = 2,
#                  hetero = TRUE, diff = TRUE, B = 500)
# print(ind)
```

```{r indep-perm}
# Sensitivity of the conclusion to the HSIC inference method
# dda.indep(m, pred = "x", data = dat, hsic.method = "permutation", B = 500)
```

## Conditional analysis

```{r cdda, fig.cap = "Conditional dCor differences with bootstrap confidence bands. Values above zero favor the target direction; a band covering zero is undecided at that moderator value."}
# mi <- lm(y ~ x * z, data = dat)
# ci <- cdda.indep(mi, pred = "x", mod = "z", modval = "JN",
#                  data = dat, diff = TRUE, hetero = TRUE, B = 500)
# summary(ci, hsic.diff = TRUE, dcor.diff = TRUE)
# plot(ci, stat = "dcor.diff")
```

## Robustness and decision stability

```{r robust-bag}
# ind_rob <- dda.indep(m, pred = "x", data = dat, robust = TRUE,
#                      hetero = TRUE, B = 500)
# bag <- dda.bagging(ind, data = dat, iter = 200, agg_stat = "trimmed")
# summary(bag, show = c("hsic", "dcor", "bp"))   # TABLE 2
```

## Synthesis

# Practical Guidance

<!-- Sample size; how much non-normality is enough; what to do when criteria
     disagree (give an explicit ordering and defend it); behavior under
     confounding; choosing hsic.method / boot.type / B; a reporting checklist;
     common errors. -->

# Discussion

<!-- What was established; limitations of the method; limitations of the
     software; where DDA sits alongside experimental design; planned
     extensions. -->

# Open Practices Statement

```{r session, echo = FALSE}
sessionInfo()
```

<!-- Package version and source; R version and platform; OSF DOI for scripts and
     data; random seeds. -->

# References

<div id="refs"></div>

\appendix

# Full reproducible script

# Simulation sanity check (optional)
