---
title: "Package 'dda'"
output:
  pdf_document:
    toc: true
    toc_depth: '4'
  html_document:
    theme: spacelab
    highlight: kate
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    toc_depth: 4
    fig_width: 14
    fig_height: 9
    fig_caption: true
    number_sections: true
vignette: "%\\VignetteIndexEntry{Introduction to Direction Dependence Analysis (dda)} %\\VignetteEngine{knitr::rmarkdown} %\\VignetteEncoding{UTF-8}\n"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This is a vignette for the `dda` package. It provides an overview of the package and its functionality. `dda` is a package based on the Direction Dependence Analysis (DDA) framework proposed by Wiedermann, W., & von Eye, A (2015)](https://doi.org/10.1177/0165025415582056). The package compares competing directional linear models: a) the target model (initially specified) `x -> y` and b) the reversed model `y -> x` (reverse specification). The package provides a set of functions to perform the DDA analysis, to test and to visualize the results.

# Installation

You can install the pre-released version of `dda` from GitHub with: 

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("wwiedermann/dda")
library(dda)
```

```{r, message = FALSE, warning = FALSE, include=FALSE, echo = FALSE}
devtools::install_github("hirnia/dda")
library(dda)
```

# Implementation Example: One Simulation in Many Renditions

## 0: Data Generation

### Data Generation

```{r}
set.seed(123)

n <- 1000

## --- generate moderator

z <- sort(rnorm(n))
z1 <- z[z <= 0]
z2 <- z[z > 0]
z.categ <- as.factor(ifelse(z < 0, 0, 1))

## --- x -> y when m <= 0

x1 <- rchisq(length(z1), df = 4) - 4
e1 <- rchisq(length(z1), df = 3) - 3
y1 <- 0.5 * x1 + e1

## --- y -> x when m > 0

y2 <- rchisq(length(z2), df = 4) - 4
e2 <- rchisq(length(z2), df = 3) - 3
x2 <- 0.25 * y2 + e2

y <- c(y1, y2)
x <- c(x1, x2)
dat <- data.frame(x,y,z,z.categ)

m <- lm(y ~ x*z, data = dat)

summary(m)
#plot(y ~ x, data = dat, pch = 21, bg = as.factor(z <= 0))
```

# 1. Independence Properties (`indep`)

## 1.A `dda.indep` implementation

### 1.A.1 Continuous Predictor Example

#q:Error: 'dda.indep' is not an exported object from 'namespace:dda', why
#a: The function is not exported, so you need to use the `::` operator to access it.
#q: excellent, but I am using the `dda` package, so why is the function not exported?
#a: The function is not exported because it is not intended to be used by the end-user. It is used internally by other functions in the package.

```{r TEST}
m.categ <- lm(y ~ x*z.categ, data = dat)

factor.indep <- cdda.indep(m.categ, pred = "x", mod = "z.categ", 
                          modval = "factor", data = dat, boot.type = "bca",
                          nlfun = 2, diff = TRUE, hetero = TRUE, B = 1000)
```



```{r}
dda.indep(y ~ x + z, pred = "x", data = dat, B = 1000)
```

### 1.A.2 Categorical Predictor Example
```{r}
dda.indep(y ~ x + z.categ, pred = "x", data = dat, B = 1000)
```

## 1.B `cdda.indep` implementation

### 1.B.1 Factor Level

```{r}
m.categ <- lm(y ~ x*z.categ, data = dat)

factor.indep <- cdda.indep(m.categ, pred = "x", mod = "z.categ", 
                          modval = "factor", data = dat, boot.type = "bca",
                          nlfun = 2, diff = TRUE, hetero = TRUE, B = 1000)
```

```{r}
factor.indep
```

```{r}
summary(factor.indep, hsic.diff = TRUE, dcor.diff = TRUE,
        mi.diff = TRUE, nlfun = T)
```

```{r}
plot.cddaindep(factor.indep, stat = "hsic.diff")
plot.cddaindep(factor.indep, stat = "dcor.diff")
plot.cddaindep(factor.indep, stat = "mi.diff")
```


### 1.B.2 Pick A Point

The poster looks good. Would it make sense to present some other independence tests than Breusch-Pagan that show clearer results in favor of the target model? Similarly, maybe you find another example of a plot than kurtosis difference with clearer results. Other than that, I don't think the poster needs much revision.

```{r}
point.indep <- cdda.indep(m, pred = "x", mod = "z", data = dat, nlfun = 2,
                          modval = c(-1, 0, 1), diff = TRUE, hetero = TRUE, point = 0)
```

```{r}
# point.indep
summary(point.indep, hsic.diff = T, dcor.diff = T, mi.diff = T)
```

```{r}
plot.cddaindep(point.indep, stat = "hsic.diff")
plot.cddaindep(point.indep, stat = "dcor.diff")
plot.cddaindep(point.indep, stat = "mi.diff")
```

### 1.B.3 Mean
```{r}
mean.indep <- cdda.indep(m, pred = "x", mod = "z", data = dat, nlfun = 2,
                          modval = "mean", diff = TRUE, hetero = TRUE)
```

```{r}
#mean.indep
summary(mean.indep, hsic.diff = TRUE, dcor.diff = TRUE, mi.diff = TRUE)
```

```{r}
plot.cddaindep(mean.indep, stat = "hsic.diff")
plot.cddaindep(mean.indep, stat = "dcor.diff")
plot.cddaindep(mean.indep, stat = "mi.diff")
```

### 1.B.4 Median
```{r}
median.indep <- cdda.indep(m, pred = "x", mod = "z", data = dat, nlfun = 2,
                          modval = "median", diff = TRUE, hetero = TRUE, point = 0)
```

```{r}
#median.indep
summary(median.indep, hsic.diff = TRUE, dcor.diff = TRUE, mi.diff = TRUE)
```

```{r}
plot.cddaindep(median.indep, stat = "hsic.diff")
plot.cddaindep(median.indep, stat = "dcor.diff")
plot.cddaindep(median.indep, stat = "mi.diff")
```


### 1.B.5 Johnson-Neyman, 4 Levels
```{r}
jn.indep.b <- cdda.indep(m, pred = "x", mod = "z", data = dat, nlfun = 2,
                         modval = "JN", JN.length = 4, diff = TRUE, hetero = TRUE)
```

```{r}
#jn.indep
summary(jn.indep.b, hsic.diff = TRUE, dcor.diff = TRUE, mi.diff = TRUE)
```

```{r}
plot.cddaindep(jn.indep.b, stat = "hsic.diff")
plot.cddaindep(jn.indep.b, stat = "dcor.diff")
plot.cddaindep(jn.indep.b, stat = "mi.diff")
```

# 2. Residual Distribution Properties (`resdist`)

## 2.A `dda.resdist` implementation

### 2.A.1 Continuous Predictor Example
```{r}
dda.resdist(y ~ x + z, pred = "x", data = dat, B = 1000)
```

### 2.A.2 Categorical Predictor Example
```{r}
dda.resdist(y ~ x + z.categ, pred = "x", data = dat, B = 1000)
```

# 3. Variable Distribution Properties (`vardist`)

## 3.A `dda.vardist` implementation

### 3.A.1 Continuous Predictor Example
```{r}
dda.vardist(y ~ x + z, pred = "x", data = dat, B = 1000)
```

### 3.A.2 Categorical Predictor Example
```{r}
dda.vardist(y ~ x + z.categ, pred = "x", data = dat, B = 1000)
```

## 3.B `cdda.vardist` implementation

### 3.B.1 Factor Level

```{r}
factor.vardist <- cdda.vardist(m.categ, pred = "x", mod = "z.categ",
                               modval = "factor", data = dat)
```

```{r}
factor.vardist
```

```{r}
summary(factor.vardist, coskew = TRUE, cokurt = TRUE)
```

```{r}
plot(factor.vardist, stat = "rhs", ylim = c(-0.3, 0.3))
plot(factor.vardist, stat = "cokurt", ylim = c(-10, 30))
plot(factor.vardist, stat = "rcc", ylim = c(-5, 15))
plot(factor.vardist, stat = "rtanh", ylim = c(0, 0.03))
```

## 3.B.2 Pick A Point

```{r}
point.vardist <- cdda.vardist(m, pred = "x", mod = "z", data = dat,
                          modval = c(-1, 0, 1), point = 0)
```

```{r}
#point.vardist
summary(point.vardist, coskew = TRUE, cokurt = TRUE)
```

## 3.B.3 Mean

```{r}
mean.vardist <- cdda.vardist(m, pred = "x", mod = "z",
                             data = dat, modval = "mean")
```

```{r}
summary(mean.vardist, coskew = TRUE, cokurt = TRUE)
```

```{r}
plot(mean.vardist, stat = "cokurt", ylim = c(-10, 40))
plot(mean.vardist, stat = "rcc", ylim = c(-5, 25))
plot(mean.vardist, stat = "rtanh", ylim = c(-0.01, 0.03))
```

## 3.B.4 Median 

```{r}
median.vardist <- cdda.vardist(m, pred = "x", mod = "z",
                               data = dat, modval = "median")
```

```{r}
#median.vardist
summary(median.vardist, coskew = TRUE, cokurt = TRUE)
```

```{r}
plot(median.vardist, stat = "rhs", ylim = c(-0.2, 0.3))
plot(median.vardist, stat = "cokurt", ylim = c(-10, 35))
plot(median.vardist, stat = "rcc", ylim = c(-5, 25))
plot(median.vardist, stat = "rtanh", ylim = c(-0.01, 0.03))
```

## 3.B.5 Johnson-Neyman, 4 Levels

```{r}
jn.vardist.b <- cdda.vardist(m, pred = "x", mod = "z",
                           data = dat, modval = "JN", JN.length = 4)
```

```{r}
#jn.vardist
summary(jn.vardist.b, coskew = TRUE, cokurt = TRUE)
```

```{r}
plot(jn.vardist.b, stat = "rhs", ylim = c(-0.05, 0.05))
plot(jn.vardist.b, stat = "cokurt", ylim = c(-2, 5))
plot(jn.vardist.b, stat = "rcc", ylim = c(-2, 5))
plot(jn.vardist.b, stat = "rtanh", ylim = c(-0.005, 0.005))
```



