---
title: "Package 'dda'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
can_decrypt <- gargle:::secret_can_decrypt("cdda")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE,
  eval = can_decrypt
)
```

```{r eval = !can_decrypt, echo = FALSE, comment = NA}
#message("No token available. Code chunks will not be evaluated.")
```

# Introduction

This is a vignette for the `dda` package. It provides an overview of the package and its functionality. `dda` is a package based on the Direction Dependence Analysis (DDA) framework proposed by Wiedermann, W., & von Eye, A (2015)](https://doi.org/10.1177/0165025415582056). The package compares competing directional linear models: a) the target model (initially specified) `x -> y` and b) the reversed model `y -> x` (reverse specification). The package provides a set of functions to perform the DDA analysis, to test and to visualize the results.

# Installation

You can install the pre-released version of `dda` from GitHub with: 

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("wwiedermann/dda")
```

